---
title: "TimeSeries_Analysis"
author: "Yitong Chen"
date: "10/16/2020"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE}
library(ggplot2)
library(forecast)
library(fpp)
library(gridExtra)
library(astsa)
library(forecast)
library(TSA)
library(cowplot)
library(dplyr)
```

### Introduction: 
This Data Set contains 180 data points with column one representing time (in months) and column two representing some unknown observations. <br/>

load the dataset
```{r}
data1 <- read.csv("./data1.csv")
```

### 1 Exploratory Data Analysis
#### 1.1 Stationarity & Variance
Let's define the our initial Time Series as $A_{t}$. <br/><br/>
We began our EDA by first observing its general trend and shape (Figure 1.1a). Through its obvious upward trend, heteroscedasticity, and potential seasonality, we understood that it was clearly not stationary time series. Thus, to address the issue, we began our journey to make predictions by conducting a variance stabilizing transformation followed by differencing in order to achieve stationary data for the purpose of using an appropriate (M)(S)ARIMA model.<br/><br/>

To remove the unstationarities, we do the following transformations: <br/>
1) Take Logarithm        ($A_{t} -> L_{t}$)    : $L_{t} = log{A_t}$ <br/>
2) Take First Difference ($L_{t} -> X_{t}$)    : $X_{t} = L_{t} - L_{t-1}$
```{r, warning=FALSE}
# figure 1.1a
plot1 <- ggplot() +
  geom_line(data = data1, aes(X, x), color = "black")  + labs(title="Plot of Orginal Time Series, At",
        x ="Time", y = "X")

# figure 1.1b  take logarithm
data1 <- data1 %>% mutate(Lt = log(data1$x))
plot2 <- ggplot() +
  geom_line(data = data1, aes(X, Lt), color = "black")  + labs(title="Plot of Log Transformed Time Series, Lt",
        x ="Time", y = "Nt")

# figure 1.1c  take 1st difference
data1 <- data1 %>% mutate(Xt = c(NA, diff(data1$x, lag = 1)))
plot3 <- ggplot() +
  geom_line(data = data1, aes(X, Xt), color = "black") + geom_hline(yintercept = 0, 
                color = "blue", size=0.7) + labs(title="Plot of 1st Difference Time Series",
        x ="Time", y = "Xt")

grid.arrange(plot1, plot2, plot3, ncol = 2)
```


1.2 ACF & PACF of Xt
Looking at the ACF and PACF of the Xt(Figure 1.2a), we observe that both have a slight tapering off behavior in small lags, implying a non-seasonal ARMA for Xt (we’ll discuss this in greater detail in our model selection process from Section 3); at big lags, both ACF and PACF also show a tapering off behavior in a seasonal pattern. Such a behavior directs us to utilizing Frequency Domain Analysis to detect the exact seasonality (this will be done in Section 2).
```{r}
p1 <- ggAcf(data1$Xt, lag.max = 120)+ ggtitle("1st diff of sqrt original")
p2 <- ggPacf(data1$Xt, lag.max = 120) + ggtitle("1st diff of sqrt original")
grid.arrange(p1, p2, nrow = 2)
```