---
title: "TimeSeries_Analysis"
author: "Yitong Chen"
date: "10/16/2020"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE}
library(ggplot2)
library(forecast)
library(fpp)
library(gridExtra)
library(astsa)
library(forecast)
library(TSA)
library(cowplot)
library(dplyr)
```

### Introduction: 
This Data Set contains 180 data points with column one representing time (in months) and column two representing some unknown observations. <br/>

load the dataset
```{r}
data1 <- read.csv("./data1.csv")
```

### 1 Exploratory Data Analysis
#### 1.1 Stationarity & Variance
Let's define the our initial Time Series as $A_{t}$. <br/><br/>
We began our EDA by first observing its general trend and shape (Figure 1.1a). Through its obvious upward trend, heteroscedasticity, and potential seasonality, we understood that it was clearly not stationary time series. Thus, to address the issue, we began our journey to make predictions by conducting a variance stabilizing transformation followed by differencing in order to achieve stationary data for the purpose of using an appropriate (M)(S)ARIMA model.<br/><br/>

To remove the unstationarities, we do the following transformations: <br/>
1) Take Logarithm        ($A_{t} -> L_{t}$)    : $L_{t} = log{A_t}$ <br/>
2) Take First Difference ($L_{t} -> X_{t}$)    : $X_{t} = L_{t} - L_{t-1}$
```{r, warning=FALSE}
# figure 1.1a
plot1 <- ggplot() +
  geom_line(data = data1, aes(X, x), color = "black")  + labs(title="Plot of Orginal Time Series, At",
        x ="Time", y = "X")

# figure 1.1b  take logarithm
data1 <- data1 %>% mutate(Lt = log(data1$x))
plot2 <- ggplot() +
  geom_line(data = data1, aes(X, Lt), color = "black")  + labs(title="Plot of Log Transformed Time Series, Lt",
        x ="Time", y = "Nt")

# figure 1.1c  take 1st difference
data1 <- data1 %>% mutate(Xt = c(NA, diff(data1$x, lag = 1)))
plot3 <- ggplot() +
  geom_line(data = data1, aes(X, Xt), color = "black") + geom_hline(yintercept = 0, 
                color = "blue", size=0.7) + labs(title="Plot of 1st Difference Time Series",
        x ="Time", y = "Xt")

grid.arrange(plot1, plot2, plot3, ncol = 2)
```


1.2 ACF & PACF of Xt
Looking at Figure 1.2a, we observe that the ACF and PACF of $X_{t}$ are significant at around lag 6, 12, 18, 24, a.k.a implying a seasonal pattern of 6 periods. This is suggesting us to remove the seasonalities by differencing.
```{r}
p1 <- ggAcf(data1$Xt, lag.max = 120)+ ggtitle("1st diff of sqrt original")
p2 <- ggPacf(data1$Xt, lag.max = 120) + ggtitle("1st diff of sqrt original")
grid.arrange(p1, p2, nrow = 2)
```

1.3 Seasonal differencing
Take Sixth Difference on $X_{t}$ ($X_{t} -> Y_{t}$)    : $Y_{t} = X_{t} - X_{t-6}$
```{r, warning=FALSE}
data1 <- data1 %>% mutate(Yt = c(NA, NA,NA,NA,NA, diff(data1$Xt, lag = 5)))
plot4 <- ggplot() +
  geom_line(data = data1, aes(X, Yt), color = "black") + geom_hline(yintercept = 0, 
                color = "blue", size=0.7) + labs(title="Plot of Yt Time Series with Seasonality Removed",
        x ="Time", y = "Yt")
plot4
```


```{r}
adf.test(na.remove(data1$Yt))
```


1.4 select a model
Augmented Dicky-Fuller test indicates if the data correponds to a stationary process, and the above test shows $Y_{t}$ is significant at 0.01 confidence level. Therefore, we are quite convinced that $Y_{t}$ is a stationary process. We'll fit an ARMA model from here.<br/>
From PACF, we see an obvious 
```{r}
p3 <- ggAcf(data1$Yt, lag.max = 60)+ ggtitle("Yt ACF")
p4 <- ggPacf(data1$Yt, lag.max = 60) + ggtitle("Yt PACF")
grid.arrange(p3, p4, nrow = 2)
```



